---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE
)
```

[![Travis-CI Build Status](https://travis-ci.org/raymondben/peranavolley.svg?branch=master)](https://travis-ci.org/raymondben/peranavolley)

# peranavolley

An R package for reading [Perana Sports VBStats](http://peranasports.com/software/vbstatshd/) volleyball scouting files as well as [Perana Sports VideoTagger](http://peranasports.com/software/videotagger/) files.

See also [datavolley](https://github.com/raymondben/datavolley), which provides similar functionality for reading DataVolley files into R. The conventions (data structures, variable naming, etc) used in the peranavolley package follow those of the datavolley package.

## Installation

You can install peranavolley from GitHub with:

```{r eval = FALSE}
library(remotes)
install_github("raymondben/peranavolley")
```

## Example

Read one of the example data files bundled with the package:

```{r}
library(peranavolley)
x <- pv_read(pv_example_file())
summary(x)
```

Number of serves by team:

```{r}
serve_idx <- which(plays(x)$skill == "Serve")
table(plays(x)$team[serve_idx])
```

Attack kill rate by player, using `dplyr`:

```{r}
library(dplyr)
plays(x) %>% dplyr::filter(skill == "Attack") %>%
  group_by(team, player_name, player_id) %>%
  dplyr::summarize(`N attacks` = n(), `Kill rate` = mean(evaluation == "Winning attack", na.rm = TRUE)) %>%
  ungroup %>%
  dplyr::arrange(desc(`Kill rate`)) %>% dplyr::select(-player_id)
```
